---
title: "p8105_hw1_jz3036.Rmd"
output: github_document
author: "Junyuan Zheng (jz3036)"
date: 2018-09-13
---
Homework 1

# Problem 1

Create a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution
```{r}
samp = runif(10, min=0, max=5)
samp
```

* A logical vector indicating whether elements of the sample are greater than 2
```{r}
logi_vec = samp > 2
logi_vec
```

* A (length-10) character vector
```{r}
char_vec = c('hello', 'there', '!', 'this', 'is', 'a', 'length-10', 'character', 'vector', ':)')
char_vec
```

* A (length-10) factor vector
```{r}
fact_vec = factor(c('hello', 'there', '!', 'this', 'is', 'a', 'length-10', 'factor', 'vector', ':)'))
fact_vec
```

* Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why?

the mean() function works just fine for numeric values:
```{r}
mean(samp)
```
    
the mean() function even works for logical vectors, the result reflect the proportion of True in the vector:
    
```{r}
mean(logi_vec)
```

the mean() function could not work for charactor vector, since it makes no sense to count mean of charactors:
```{r}
mean(char_vec)
```

similarly, the mean() function could not work for charactor vector:
```{r}
mean(fact_vec)
```

* try using as.numeric() function to the logical, character, and factor variables
```{r eval = FALSE}
as.numeric(logi_vec)
as.numeric(char_vec)
as.numeric(fact_vec)
```

