---
title: "p8105_hw1_jz3036.Rmd"
output: github_document
author: "Junyuan Zheng (jz3036)"
date: 2018-09-13
---
Homework 1

# Problem 1

Create a data frame comprised of:

* A random sample of size 10 from a uniform[0, 5] distribution
```{r}
samp = runif(10, min=0, max=5)
samp
```

* A logical vector indicating whether elements of the sample are greater than 2
```{r}
logi_vec = samp > 2
logi_vec
```

* A (length-10) character vector
```{r}
char_vec = c('hello', 'there', '!', 'this', 'is', 'a', 'length-10', 'character', 'vector', ':)')
char_vec
```

* A (length-10) factor vector
```{r}
fact_vec = factor(c('hello', 'there', '!', 'this', 'is', 'a', 'length-10', 'factor', 'vector', ':)'))
fact_vec
```

* Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why?

the mean() function works just fine for numeric values:
```{r}
mean(samp)
```
    
the mean() function even works for logical vectors, the result reflect the proportion of True in the vector:
    
```{r}
mean(logi_vec)
```

the mean() function could not work for charactor vector and returned Missing Value, since it makes no sense to count mean of charactors:
```{r}
mean(char_vec)
```

similarly, the mean() function could not work for charactor vector and returned Missing Value:
```{r}
mean(fact_vec)
```

* As shown in the chunk below, this is to try using as.numeric() function to the logical, character, and factor variables.
    * For logic vector, as.numeric() function converts "True" into 1, "False" into 0.
    * For character vector, since as.numeric() function doesn't know how to convert those in to number, it's forced to return "NA" as missing value.
    * for factor vector, as.numeric() function concerts it into numbers representing the level of each character in the vector.
```{r eval = FALSE}
as.numeric(logi_vec)
as.numeric(char_vec)
as.numeric(fact_vec)
```

* Convert your character variable from character to factor to numeric. What happens?
    * as.factor() function successfully converted character vector into factor, and automatically set the level to each character.
    * then the as.numeric() function converted the new factor vector into numeric numbers according to the level of each character.
```{r}
new_fact_vec = as.factor(char_vec)
new_fact_vec
as.numeric(new_fact_vec)
```

* convert your factor variable from factor to character to numeric. What happens?
    * as.character() function successfully concerted factor vertor into character vector.
    * While as.numeric() function still failed and returned missing value.
```{r}
new_char_vec = as.character(fact_vec)
new_char_vec
as.numeric(new_char_vec)
```

# Problem 2

## Create a data frame comprised of:

* x: a random sample of size 1000 from a standard Normal distribution.
```{r}
x = rnorm(1000)
```

* y: a random sample of size 1000 from a standard Normal distribution.
```{r}
y = rnorm(1000)
```

* A logical vector indicating whether the x + y > 0
```{r}
logi_vec = x + y > 0
```

* A numeric vector created by coercing the above logical vector
```{r}
nume_vec = as.numeric(logi_vec)
```

* A factor vector created by coercing the above logical vector
```{r}
fact_vec = as.factor(logi_vec)
```

* Vector x contains `r length(x)` numbers randomly generated from a standard Normal distribution. The mean of x is `r mean(x)`, and the median is `r median(x)`. The proportion of the 'TRUE' in the factor vector is `r table(fact_vec)['TRUE']/length(x)`.

## Make a scatterplot of y vs x; color points using the logical variable
```{r}
library(tidyverse)
plot_df = tibble(x = x, y = y)
ggplot(plot_df, aes(x = x, y = y, color = logi_vec)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

## Make a scatterplot of y vs x; color points using the numeric variable
```{r}
library(tidyverse)
plot_df = tibble(x = x, y = y)
ggplot(plot_df, aes(x = x, y = y, color = nume_vec)) + geom_point()
```

## Make a scatterplot of y vs x; color points using the factor variable
```{r}
library(tidyverse)
plot_df = tibble(x = x, y = y)
ggplot(plot_df, aes(x = x, y = y, color = fact_vec)) + geom_point()
```